Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae

### 2023-04-03 paper/markdown/figures_and_tables.Rmd
Format of figures and references im Text

bibliography: hyposmia_biblio.bib
csl: frontiers_medical.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

@@

```{r fig-consort, fig.width = mm_inch(figures$consort$w), fig.height = mm_inch(figures$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.'}

figures$consort$plot

__Figure \@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.__

@@

_(B) Percentages of individuals with smell and taste disorders in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset. Numbers of complete observations are indicated under the plots._

@@

```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.'}

@@

__Figure \@ref(fig:fig-sympt-mds). Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.__

@@

@@

```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.'}

@@

__Figure \@ref(fig:fig-apriori). Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.__

_Frequent combinations of symptoms at $\geq$ 28 days (A) and $\geq$ 3 months (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm. Symptom combination support and confidence are presented in the plots. Point color codes for lift statistic values. The hypo/anosmia and hypo/ageusia combinations are labeled._

@@

```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders defines the COVID-19 recovery clusters.'}

@@

__Figure \@ref(fig:fig-sympt-clusters). Differing duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders defines the COVID-19 recovery clusters.__

_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Mean recovery times in the recovery clusters are presented as lines, 2 $\times$ SEM intervals are displayed as tinted regions. Numbers of individuals assigned to the recovery clusters are indicated under the plots._

_Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._

@@

_Differences in minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life at the time of survey completion in the recovery clusters (STDR: smell and taste disorder recovery, RR: rapid recovery, SR: slow recovery) in the Austria (AT) and Italy (IT) survey study cohorts. Dichotomous items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization. Statistical significance for differences between the clusters was assessed by Kruskal-Wallis (numeric readouts) or $\chi^2$ test (dichotomous readouts). P values were corrected for multiple testing with Benjamini-Hochberg method. Lines represent mean values, 2 $\times$ SEM intervals are displayed as tinted regions. Numbers of individuals assigned to the recovery clusters are indicated under the plots._

>>>>

bibliography: hyposmia_biblio.bib
csl: european-archives-of-oto-rhino-laryngology.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

@@

@@

_(B) Percentages of individuals with self-reported olfactory dysfunction and taste disorders in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset. Numbers of complete observations are indicated under the plots._

_OD: self-reported olfactory dysfunction, Imp. concentration: impaired concentration, Dim. appetite: diminished appetite, Imp. walk: impaired walk, Imp. FMS: impaired fine-motor skills._

@@

```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported olfactory dysfunction and taste disorders are isolated persistent symptoms of COVID-19.'}

@@

__Figure \@ref(fig:fig-sympt-mds). Self-reported olfactory dysfunction and taste disorders are isolated persistent symptoms of COVID-19.__

@@

_OD: self-reported olfactory dysfunction, Imp. concentration: impaired concentration._

@@

```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.'}

@@

__Figure \@ref(fig:fig-apriori). Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.__

_Frequent combinations (>15% of cohort) of self-reported olfactory dysfunction (OD) and other symptoms at $\geq$ 28 days (A) and $\geq$ 3 months (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm. Symptom combination frequency (support statistic) and co-occurrence (support statistic) are presented in bubble plots. Point size and color corresponds to co-occurrence, points are labeled with percentages of co-occurrence. Imp. concentration: impaired concentration._

@@

```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, olfactory dysfunction and taste disorders defines the COVID-19 recovery clusters.'}

@@

__Figure \@ref(fig:fig-sympt-clusters). Differing duration of neurocognitive and respiratory symptoms, fatigue, olfactory dysfunction and taste disorders defines the COVID-19 recovery clusters.__

_Clustering of the survey study participants in respect to symptom-specific recovery times was done by semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Mean recovery times in the recovery clusters are presented as lines, 2 $\times$ SEM intervals are displayed as tinted regions. Numbers of individuals assigned to the recovery clusters are indicated under the plots._

_OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._

@@

_Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life at the time of survey completion in the clusters in the Austria (AT) and Italy (IT) survey study cohorts are presented. Dichotomous items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization. Statistical significance for differences between the clusters was assessed by Kruskal-Wallis (numeric readouts) or $\chi^2$ test (dichotomous readouts). P values were corrected for multiple testing with Benjamini-Hochberg method. Lines represent mean values, 2 $\times$ SEM intervals are displayed as tinted regions. Numbers of individuals assigned to the recovery clusters are indicated under the plots._

### 2023-04-03 paper/markdown/supplementary_material.Rmd
Format of figures and references im Text

---
title: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
subtitle: "Supplementary Material"
author: "Hralth after COVID-19 in Tyrol and CovILD study teams"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

@@

The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the Health after COVID-19 in Tyrol survey study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the CovILD cohort.

@@

A total of `r length(globals$hact_symptoms)` was surveyed in the survey study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__). The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to 3 months: 30 moths, up to 6 months: 3 months, over 6 months: 3 months. Acute symptoms were defined as complaints present during the first 14 days after clinical onset of COVID-19.

In the CovILD study, a total of `r length(globals$covild_symptoms)` symptoms (reduced physical performance, hyposmia/anosmia, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were prospectively recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__). Acute COVID-19 symptoms were assessed retrospectively [@Sonnweber2020; @Sonnweber2022].

@@

### Rating of hyposmia with sniffing stick test

Objective hyposmia at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item sniffing stick test as described [@Rass2021]. Clinically relevant hyposmia was defined as $\leq$ 12 correct answers [@Rass2021; @Hummel2007]. In the analysis, participants with the complete answers concerning self-reported hyposmia and complete test results were included.

@@

Data transformation and statistical analysis was done with R version 4.0.5 with _tidyverse_ data science environment [@Wickham2019]. Analysis results were visualized with _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019] as well as in-house developed _ExDA_ (https://github.com/PiotrTymoszuk/ExDA) and _figur_ (https://github.com/PiotrTymoszuk/figur) packages.

@@

Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile-quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test by Wilcoxon r effect size statistic (two groups) or Kruskal-Wallis test with $\eta^2$ effect size statistic. Differences in frequency distribution for categorical outcome variables were assessed by $\chi^2$ test with Cramer V effect size statistic. Interrater assessment of self-reported and sniffing test hyposmia was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969]. $\kappa$ significance ($\kappa\neq0$) was estimated with Wald Z test. P values were adjusted for multiple testing with Benjamini-Hochberg method [@Benjamini1995] separately for each analysis task and cohort. 
Packages _rstatix_ [@Kassambara2021], _vcd_ [@Meyer2021] and _Exda_ (https://github.com/PiotrTymoszuk/ExDA) were used for statistical hypothesis testing.

@@

## Symptom-symptom distances and multi-dimensional scaling

@@

Frequent combinations of symptoms during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.1, 2 - 10 item transaction length, confidence > 0.8 and lift > 2. The support statistic were used to estimate the symptom combination frequency. The confidence value was treated as an estimate of conditional probability of the symptom co-occurrence. The lift statistic was interpreted as a measure of the symptom dependence (lift = 1, symptoms are independent).

@@

```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (AT) cohort.'}

@@

@@

_Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._

@@

```{r fig-covild-recovery, fig.width = mm_inch(suppl$covild_recovery$w), fig.height = mm_inch(suppl$covild_recovery$h), fig.cap = 'Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

@@

__Supplementary Figure S\@ref(fig:fig-covild-recovery). Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

_Percentages of individuals with the complete longitudinal data set suffering from smell disorders (A), reduced physical performance (B) and dyspnea (C) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups. Numbers of complete observations are indicated under the plots._

@@

```{r fig-hyposmia-rater100, fig.width = mm_inch(suppl$hyposmia_rater100$w), fig.height = mm_inch(suppl$hyposmia_rater100$h), fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

@@

__Supplementary Figure S\@ref(fig:fig-hyposmia-rater100). Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

_Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic. Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method. Percentages of individuals with self-reported and test hyposmia within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots. Point size and color represents the percentage. $\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._

@@

```{r fig-hyposmia-rater360, fig.width = mm_inch(suppl$hyposmia_rater360$w), fig.height = mm_inch(suppl$hyposmia_rater360$h), fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

@@

__Supplementary Figure S\@ref(fig:fig-hyposmia-rater360). Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

_Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic. Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method. Percentages of individuals with self-reported and test hyposmia within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots. Point size and color represents the percentage. $\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._

@@

__Supplementary Figure S\@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.__
_Symptom data for acute COVID-19 (first 14 days after clinical onset) in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms. MDS coordinates are presented in point plots. Selected data points are labeled with the symptom names. Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes. Numbers of complete observations are indicated under the plots._

@@

_(C) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random._

@@

```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of COVID-19 symptoms in the survey study recovery clusters.'}

@@

__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study recovery clusters.__

_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Differences in numbers of symptoms in the first 14 days (A) and at $\geq$ 28 days (B) after clinical onset between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic. P values were corrected for multiple testing with Benjamini-Hochberg method. Symptom counts are presented in violin plots. Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges. Effect size statistics and p values are indicated in the plot caption. Numbers of complete observations are displayed under the plots._

@@

__Figure \@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__

_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Differences in sex distribution (A), frequency of comorbidity (B) and daily medication (C) between the recovery clusters were assessed by $\chi^2$ test with Cramer V effect size statistic. P values were corrected for multiple testing with Benjamini-Hochberg method. The frequencies are presented as bar plots. Effect size statistics and p values are indicated in the plot caption. Numbers of complete observations are displayed under the plots._

>>>>

---
title: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
subtitle: "Supplementary Material"
author: "Health after COVID-19 in Tyrol and CovILD study teams"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

@@

The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the Health after COVID-19 in Tyrol survey study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the observatory CovILD cohort.

@@

A total of self-reported `r length(globals$hact_symptoms)` symptoms were recorded in the survey study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__). The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to or greater than 3 months: 3 months. Acute symptoms were defined as complaints present during the first 14 days after clinical onset of COVID-19.

In the observatory CovILD cohort, a total of self-reported `r length(globals$covild_symptoms)` symptoms (reduced physical performance, olfactory dysfunction, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__). Acute COVID-19 symptoms were assessed retrospectively at the 60-day follow-up [@Sonnweber2020; @Sonnweber2022].

@@

### Rating of olfactory dysfunction with sniffing stick test

Objective olfactory dysfunction at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item sniffing stick test as described [@Rass2021]. Clinically relevant olfactory dysfunction was defined as $\leq$ 12 correct answers [@Rass2021; @Hummel2007]. In the analysis, participants with the complete answers concerning self-reported olfactory dysfunction and complete test results were included.

@@

Data transformation and statistical analysis was done with R version 4.2.0 with _tidyverse_ data science environment [@Wickham2019]. Analysis results were visualized with _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019] as well as in-house developed _ExDA_ (https://github.com/PiotrTymoszuk/ExDA) and _figur_ (https://github.com/PiotrTymoszuk/figur) packages.

@@

Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile-quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test with r effect size statistic (two groups) or Kruskal-Wallis test with $\eta^2$ effect size statistic. Differences in frequency distribution for categorical outcome variables were assessed by $\chi^2$ test with Cramer V effect size statistic. Inter-rater assessment of self-reported and sniffing test olfactory dysfunction was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969]. $\kappa$ significance ($\kappa\neq0$) was estimated with Wald Z test. P values were adjusted for multiple testing with Benjamini-Hochberg method [@Benjamini1995] separately for each analysis task and cohort. 
Packages _rstatix_ [@Kassambara2021], _vcd_ [@Meyer2021] and _Exda_ (https://github.com/PiotrTymoszuk/ExDA) were used for statistical hypothesis testing.

@@

### Symptom-symptom distances and multi-dimensional scaling

@@

Frequent combinations of symptoms during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.15, 2 - 10 item transaction length, confidence > 0.6 and lift > 2. The support statistic were used to estimate the symptom combination frequency. The confidence value was treated as an estimate of conditional probability of the symptom co-occurrence. The lift statistic was interpreted as a measure of the symptom dependence (lift = 1, symptoms are independent).

@@

```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (IT) cohort.'}

@@

```{r fig-consort, fig.width = mm_inch(suppl$consort$w), fig.height = mm_inch(suppl$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.'}

suppl$consort$plot

__Supplementary Figure S\@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.__

@@

_OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._

@@

```{r fig-covild-recovery, fig.width = mm_inch(suppl$covild_recovery$w), fig.height = mm_inch(suppl$covild_recovery$h), fig.cap = 'Kinetic of recovery from olfactory dysfunction, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

@@

__Supplementary Figure S\@ref(fig:fig-covild-recovery). Kinetic of recovery from olfactory dysfunction, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

_Percentages of individuals with the complete longitudinal data set suffering from olfactory dysfunction (OD) (A), reduced physical performance (B) and dyspnea (C) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups. Numbers of complete observations are indicated under the plots._

@@

```{r fig-hyposmia-rater100, fig.width = mm_inch(suppl$hyposmia_rater100$w), fig.height = mm_inch(suppl$hyposmia_rater100$h), fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

@@

__Supplementary Figure S\@ref(fig:fig-hyposmia-rater100). Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

_Association of self-reported and sniffing stick test rates of olfactory dysfunction (OD) was investigated with Cohen's $\kappa$ statistic. Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method. Percentages of individuals with self-reported and test OD within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots. Point size and color represents the percentage. $\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._

@@

```{r fig-hyposmia-rater360, fig.width = mm_inch(suppl$hyposmia_rater360$w), fig.height = mm_inch(suppl$hyposmia_rater360$h), fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

@@

__Supplementary Figure S\@ref(fig:fig-hyposmia-rater360). Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

_Association of self-reported and sniffing stick test rates of olfactory dysfunction (OD) was investigated with Cohen's $\kappa$ statistic. Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method. Percentages of individuals with self-reported and test OD within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots. Point size and color represents the percentage. $\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._

@@

__Supplementary Figure S\@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.__
_Symptom data for acute COVID-19 (first 14 days after clinical onset) in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms. MDS coordinates are presented in point plots. Selected data points are labeled with the symptom names. Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes. Numbers of complete observations are indicated under the plots. OD: self-reported olfactory dysfunction._

@@

_(C) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random. Imp. concentration: impaired concentration, OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._

@@

```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of COVID-19 symptoms in the survey study clusters.'}

@@

__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study clusters.__

_Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Differences in numbers of symptoms in the first 14 days (A) and at $\geq$ 28 days (B) after clinical onset between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic. P values were corrected for multiple testing with Benjamini-Hochberg method. Symptom counts are presented in violin plots. Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges. Effect size statistics and p values are indicated in the plot caption. Numbers of complete observations are displayed under the plots._

@@

__Supplementary Figure S\@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__

_Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]). Differences in sex distribution (A), frequency of comorbidity (B) and daily medication (C) between the recovery clusters were assessed by $\chi^2$ test with Cramer V effect size statistic. P values were corrected for multiple testing with Benjamini-Hochberg method. The frequencies are presented as bar plots. Effect size statistics and p values are indicated in the plot caption. Numbers of complete observations are displayed under the plots._
